\begin{eabstract}
Automated theorem proving plays a crucial role in modern program
verification techniques.  In particular, in the context of software
model checking, the interpolation problem and its generalization
called \emph{Horn clause solving} have attracted special attention.
The latter problem asks for a solution for a set of acyclic Horn
clauses containing unknown predicate variables.  The solution, i.e., a
substitution for predicate variables, then serves as a candidate of
program invariants.  However, existing methods are unsatisfactory in
that they tend to generate too specific a solution, which does not
correspond to program invariants.

To address the problem above, we propose a new algorithm for solving
Horn clauses consisting of linear arithmetics and unknown predicate
variables.  The new algorithm tries to find small predicates, based on
the hypothesis that program's invariants tend to be simple formulas.
To this end, our algorithm prepares a template of linear arithmetic
formulas of a fixed size as a candidate of each unknown predicate, so
that Horn clause solving is reduced to linear programming.  The
algorithm iteratively increases the size of the templates until a
solution is found.  It enables us to construct relatively small
solutions in a reasonable time compared to existing algorithms.  We
have implemented and incorporated our Horn clause solver into MoCHi, a
software model checker for functional programs.  We have evaluated the
effectiveness of our approach through experiments.
\end{eabstract}
